---
title: "Script for Generating Numbers Report for Selected Day"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
## Initial Setup chunk. This chunk sets some general global settings and installs and attaches the required packages

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = 'A:/BioRad CFX Processing/BioRad Generated Files/Reports')

# List of required packages
packages <- c("data.table", "tidyverse", "kableExtra")

# Check if packages are installed; install if not
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Attach required packages
library(data.table)
library(tidyverse)
library(kableExtra)
```

# Introduction

This script generates a numbers report based on the text files in a selected Reported to NVRL folder. It relies on the files in that folder being the correct results reported that day; if any other files reported on different days are present, these will be included in the report as well.

### Step 1

RStudio must be installed on your local computer. Speak to Dan or Conor about getting this set up.

### Step 2

Open up this file in RStudio, named Numbers.Rmd. The folder path is given below.

```{r, eval = FALSE, echo = TRUE, include = TRUE}
COVID 19 (A:)\DA\Conor\Scripts
```

### Step 3

Click 'Run' and 'Run All' in the taskbar above to the right.

```{r, include = FALSE}
## This chunk contains the function for generating the numbers report

numbers <- function(path){

# Set the folder path given as input
suppressWarnings(setwd(path))
  
# Read the exported text files in the folder into a single dataframe with sep and NA preferences set
data <- as.data.frame(rbindlist(lapply(list.files(path = ".", pattern = ".txt"), fread, sep="\t", na.strings="NaN")))

# Rename column names to remove whitespaces
colnames(data) <- c("Sample", "Position", "Channel", "Cp", "Call")


# Interpret result for each sample; algorithm parameters are set in the ifelse statement
data <- data %>% 
  
  # Remove duplicate samples
  distinct(Sample, Channel, .keep_all = T) %>%
  
  # Change dataframe from long to wide format so that each sample is on one row
  spread(Channel, Cp) %>%
  
  # Interpret result
  mutate(Call = ifelse(is.na(FAM) & is.na(CY5) & is.na(HEX), "Invalid", 
                ifelse(is.na(FAM) & is.na(CY5) & HEX > 0, "Negative", 
                ifelse(is.na(FAM) & CY5 >= 30 | FAM >= 30 & is.na(CY5), "Presumptive",
                ifelse(FAM < 30 & CY5 < 30 | FAM < 30 & is.na(CY5) | is.na(FAM) & CY5 < 30 | FAM < 30 & CY5 > 30 | FAM > 30 & CY5 < 30, "Positive", 
                ifelse(FAM > 30 & CY5 > 30, "Presumptive", NA))))))


# Count the numbers for each result

# Total samples
a <- length(unique(data$Sample))

# Total positive or presumptive
b <- sum(sum(data$Call == "Positive"), sum(data$Call == "Presumptive"))

# Total negative
c <- sum(data$Call == "Negative")

# Total invalid
d <- sum(data$Call == "Invalid")

# Percentage positive
e <- (sum(sum(data$Call == "Positive"), sum(data$Call == "Presumptive")) * 100 / sum(sum(sum(data$Call == "Positive"), sum(data$Call == "Presumptive")),sum(data$Call == "Negative"),sum(data$Call == "Invalid")))


# Create table
table <- data.frame(a, b, c, d, e)

# Set table column names
colnames(table) <- c("Total Samples", "Total Positive", "Negative", "Invalid", "Percentage Positive")

# Format table using kableExtra
kable(table, format = "html", row.names = FALSE) %>%
  kable_styling(full_width = F, position = "left")

}
```

### Step 4

This step generates the report. The function is in the form `numbers('path')`, where 'path' is the location of the folder you want to analyse. 

Type in the folder and click the 'run' button.

An example is given below.

```{r, warning = FALSE, echo = TRUE, include = TRUE}
numbers("01 January/23012021/For NVRL export")
```


#### Important Usage Note

For the sake of convenience, it is not necessary to type the entire folder path; just the path from "01 January" onwards.

Also, Windows separates folder names with a "\", while RStudio uses a "/" or else it won't work. Please ensure the right character is used. 


#### Troubleshooting

Speak to Conor or Dan with any issues.

