---
title: "Script for Generating Numbers Report for Selected Day"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
## Initial Setup chunk. This chunk sets some general global settings and installs and attaches the required packages

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = 'A:/BioRad CFX Processing/BioRad Generated Files/Reports')

# List of required packages
packages <- c("data.table", "tidyverse", "kableExtra")

# Check if packages are installed; install if not
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Attach required packages
library(data.table)
library(tidyverse)
library(kableExtra)
```

# Introduction

This script generates a numbers report based on the text files in a selected Reported to NVRL folder. It relies on the files in that folder being the correct results reported that day; if any other files reported on different days are present, for example re-uploaded samples, these will be included in the report as well. The script is designed to handle text files with results from all three assays; Altona, Euroimmun, and Idexx.

### Step 1

RStudio must be installed on your local computer. Speak to Dan or Conor about getting this set up.

### Step 2

Open up this file in RStudio, named Numbers.Rmd. The folder path is given below.

```{r, error = FALSE, eval = FALSE, echo = TRUE, include = TRUE}
COVID 19 (A:)\DA\Conor\Scripts
```

### Step 3

Click 'Run' and 'Run All' in the taskbar above to the right.

```{r, include = FALSE}
## This chunk contains the functions for generating the numbers report

## Usage!

# There are three basic functions to count the positives depending on the assay.
# Each function reads in .txt files from the specified folder that contain the name of the assay, e.g. _Altona.txt. 
# Each of the three functions are largely the same except for the interpretation conditions;
# i.e. Altona results are interpreted based on FAM and Cy5, while Euroimmun is interpreted based on FAM alone. There may also be different Cp thresholds. 
# The fourth and final function, named 'numbers', calls each of the three functions to count numbers and combines the results into one table.

  
## Call the Altona function  

altona_numbers <- function(){
  
  # Read in all .txt files in the folder that contain "Altona"
  altona_files <- list.files(path = ".", pattern = "Altona.txt")

  # Check if any Altona .txt files were read in; perform the function on any that were
  if(length(altona_files) != 0){
      
    # Read the exported text files in the folder into a single dataframe with sep and NA preferences set
    data <- as.data.frame(rbindlist(lapply(altona_files, fread, sep="\t", na.strings="NaN")))
  
    # Rename column names to remove whitespaces
    colnames(data) <- c("Sample", "Position", "Channel", "Cp", "Call")
  
  
    # Interpret result for each sample; algorithm parameters are set in the ifelse statement
    data <- data %>%
  
    # Remove duplicate samples
    distinct(Sample, Channel, .keep_all = T) %>%
  
    # Change dataframe from long to wide format so that each sample is on one row
    spread(Channel, Cp) %>%
  
    # Interpret result using dplyr::mutate and ifelse conditions; invalid, positive, or negative
    mutate(Call = ifelse(is.na(FAM) & is.na(CY5) & is.na(HEX), "Invalid", 
                    ifelse(is.na(FAM) & is.na(CY5) & HEX > 0, "Negative", 
                    ifelse(FAM > 0 | CY5 > 0, "Positive", NA))))
    
    
    # Open a new column that will be used to designate samples based on Cp range - 15-20, 30-35, etc.
    data[, "Cp_Call"] <- NA
    
    
    # Designate samples based on Cp range using dplyr::mutate and ifelse conditions
    data <- data %>%
      
      mutate(Cp_Call = ifelse(is.na(FAM) & CY5 > 35 | FAM > 35, "35-40",
                       ifelse(is.na(FAM) & CY5 > 30 | FAM > 30, "30-35", 
                       ifelse(is.na(FAM) & CY5 > 25 | FAM > 25, "25-30",
                       ifelse(is.na(FAM) & CY5 > 20 | FAM > 20, "20-25",
                       ifelse(is.na(FAM) & CY5 > 15 | FAM > 15, "15-20",
                       ifelse(is.na(FAM) & CY5 > 10 | FAM > 10, "10-15",
                       ifelse(is.na(FAM) & CY5 > 5 | FAM > 5, "5-10",
                       ifelse(is.na(FAM) & CY5 > 0 | FAM > 0, "0-5", NA)))))))))
    
    
    ## Count the numbers for each result
    
    # Total samples
    tot_samples <- length(unique(data$Sample))
    
    # Total Cp 0 to 5
    cp_5 <- sum(data$Cp_Call == "0-5", na.rm = T)
    
    # Total Cp 5 to 10
    cp_10 <- sum(data$Cp_Call == "5-10", na.rm = T)
    
    # Total Cp 10 to 15
    cp_15 <- sum(data$Cp_Call == "10-15", na.rm = T)
    
    # Total Cp 15 to 20
    cp_20 <- sum(data$Cp_Call == "15-20", na.rm = T)
    
    # Total Cp 20 to 25
    cp_25 <- sum(data$Cp_Call == "20-25", na.rm = T)
    
    # Total Cp 25 to 30
    cp_30 <- sum(data$Cp_Call == "25-30", na.rm = T)
    
    # Total Cp 30 to 35
    cp_35 <- sum(data$Cp_Call == "30-35", na.rm = T)
    
    # Total Cp 35 to 40
    cp_40 <- sum(data$Cp_Call == "35-40", na.rm = T)
    
    # Total positives
    tot_pos <- sum(data$Call == "Positive")
    
    # Total negative
    tot_neg <- sum(data$Call == "Negative")
    
    # Total invalid
    tot_inv <- sum(data$Call == "Invalid")
  
    # Percentage positive
    pc_pos <- round((tot_pos * 100 / tot_samples), 2)
    
    # Create table with Cp counts as percentages of total positives
    altona_table <- data.frame(tot_samples, tot_pos, tot_inv, pc_pos, cp_5, cp_10, cp_15, cp_20, cp_25, cp_30, cp_35, cp_40)
  
  
    # Set table column names
    colnames(altona_table) <- c("Samples", "Positives", "Invalids", "% Positive", "% 0-5", "% 5-10", "% 10-15", "% 15-20", "% 20-25", "% 25-30", "% 30-35", "% 35-40")
  
    # Handle missing values that result from getting percentage of 0 positives
    altona_table[altona_table == "NaN"] <- 0.00
    
    return(altona_table)
  }
}


## Call the Euroimmun function  

euroimmun_numbers <- function(){
  
  # Read in all .txt files in the folder that contain "Euroimmun"
  euroimmun_files <- list.files(path = ".", pattern = "Euroimmun.txt")

  # Check if any Euroimmun .txt files were read in; perform the function on any that were
  if(length(euroimmun_files) != 0){
      
    # Read the exported text files in the folder into a single dataframe with sep and NA preferences set
    data <- as.data.frame(rbindlist(lapply(euroimmun_files, fread, sep="\t", na.strings="NaN")))
  
    # Rename column names to remove whitespaces
    colnames(data) <- c("Sample", "Position", "Channel", "Cp", "Call")
  
  
    # Interpret result for each sample; algorithm parameters are set in the ifelse statement
    data <- data %>%
  
    # Remove duplicate samples
    distinct(Sample, Channel, .keep_all = T) %>%
  
    # Change dataframe from long to wide format so that each sample is on one row
    spread(Channel, Cp) %>%
  
    # Interpret result using dplyr::mutate and ifelse conditions
     mutate(Call = ifelse(is.na(FAM) & is.na(HEX), "Invalid", 
                    ifelse(is.na(FAM) & HEX > 0, "Negative", 
                    ifelse(FAM > 0, "Positive", NA))))
    
    data[, "Cp_Call"] <- NA
    
    data <- data %>%
      
      mutate(Cp_Call = ifelse(FAM > 35, "35-40",
                       ifelse(FAM > 30, "30-35", 
                       ifelse(FAM > 25, "25-30",
                       ifelse(FAM > 20, "20-25",
                       ifelse(FAM > 15, "15-20",
                       ifelse(FAM > 10, "10-15",
                       ifelse(FAM > 5, "5-10",
                       ifelse(FAM > 0, "0-5", NA)))))))))
    
    ## Count the numbers for each result
    
    # Total samples
    tot_samples <- length(unique(data$Sample))
    
    # Total Cp 0 to 5
    cp_5 <- sum(data$Cp_Call == "0-5", na.rm = T)
    
    # Total Cp 5 to 10
    cp_10 <- sum(data$Cp_Call == "5-10", na.rm = T)
    
    # Total Cp 10 to 15
    cp_15 <- sum(data$Cp_Call == "10-15", na.rm = T)
    
    # Total Cp 15 to 20
    cp_20 <- sum(data$Cp_Call == "15-20", na.rm = T)
    
    # Total Cp 20 to 25
    cp_25 <- sum(data$Cp_Call == "20-25", na.rm = T)
    
    # Total Cp 25 to 30
    cp_30 <- sum(data$Cp_Call == "25-30", na.rm = T)
    
    # Total Cp 30 to 35
    cp_35 <- sum(data$Cp_Call == "30-35", na.rm = T)
    
    # Total Cp 35 to 40
    cp_40 <- sum(data$Cp_Call == "35-40", na.rm = T)
    
    # Total positives
    tot_pos <- sum(data$Call == "Positive")
    
    # Total negative
    tot_neg <- sum(data$Call == "Negative")
    
    # Total invalid
    tot_inv <- sum(data$Call == "Invalid")
  
    # Percentage positive
    pc_pos <- round((tot_pos * 100 / tot_samples), 2)
    
    # Create table with Cp counts as percentages of total positives
    euroimmun_table <- data.frame(tot_samples, tot_pos, tot_inv, pc_pos, cp_5, cp_10, cp_15, cp_20, cp_25, cp_30, cp_35, cp_40)
  
  
    # Set table column names
    colnames(euroimmun_table) <- c("Samples", "Positives", "Invalids", "% Positive", "% 0-5", "% 5-10", "% 10-15", "% 15-20", "% 20-25", "% 25-30", "% 30-35", "% 35-40")
  
    # Handle missing values that result from getting percentage of 0 positives
    euroimmun_table[euroimmun_table == "NaN"] <- 0.00
    
    return(euroimmun_table)
  }
}


## Call the Idexx function  

idexx_numbers <- function(){
  
  # Read in all .txt files in the folder that contain "Idexx"
  idexx_files <- list.files(path = ".", pattern = "Idexx.txt")

  
  # Check if any Idexx .txt files were read in; perform the function on any that were
  if(length(idexx_files) != 0){
      
    # Read the exported text files in the folder into a single dataframe with sep and NA preferences set
    data <- as.data.frame(rbindlist(lapply(idexx_files, fread, sep="\t", na.strings="NaN")))
  
    # Rename column names to remove whitespaces
    colnames(data) <- c("Sample", "Position", "Channel", "Cp", "Call")
  
  
    # Interpret result for each sample; algorithm parameters are set in the ifelse statement
    data <- data %>%
  
    # Remove duplicate samples
    distinct(Sample, Channel, .keep_all = T) %>%
  
    # Change dataframe from long to wide format so that each sample is on one row
    spread(Channel, Cp) %>%
  
    # Interpret result using dplyr::mutate and ifelse conditions
    mutate(Call = ifelse(is.na(FAM) & is.na(HEX), "Invalid", 
                    ifelse(is.na(FAM) & HEX > 0, "Negative", 
                    ifelse(FAM > 0, "Positive", NA))))
    
    data[, "Cp_Call"] <- NA
    
    data <- data %>%
      
      mutate(Cp_Call = ifelse(FAM > 35, "35-40",
                       ifelse(FAM > 30, "30-35", 
                       ifelse(FAM > 25, "25-30",
                       ifelse(FAM > 20, "20-25",
                       ifelse(FAM > 15, "15-20",
                       ifelse(FAM > 10, "10-15",
                       ifelse(FAM > 5, "5-10",
                       ifelse(FAM > 0, "0-5", NA)))))))))
    
    ## Count the numbers for each result
    
    # Total samples
    tot_samples <- length(unique(data$Sample))
    
    # Total Cp 0 to 5
    cp_5 <- sum(data$Cp_Call == "0-5", na.rm = T)
    
    # Total Cp 5 to 10
    cp_10 <- sum(data$Cp_Call == "5-10", na.rm = T)
    
    # Total Cp 10 to 15
    cp_15 <- sum(data$Cp_Call == "10-15", na.rm = T)
    
    # Total Cp 15 to 20
    cp_20 <- sum(data$Cp_Call == "15-20", na.rm = T)
    
    # Total Cp 20 to 25
    cp_25 <- sum(data$Cp_Call == "20-25", na.rm = T)
    
    # Total Cp 25 to 30
    cp_30 <- sum(data$Cp_Call == "25-30", na.rm = T)
    
    # Total Cp 30 to 35
    cp_35 <- sum(data$Cp_Call == "30-35", na.rm = T)
    
    # Total Cp 35 to 40
    cp_40 <- sum(data$Cp_Call == "35-40", na.rm = T)
    
    # Total positives
    tot_pos <- sum(data$Call == "Positive")
    
    # Total negative
    tot_neg <- sum(data$Call == "Negative")
    
    # Total invalid
    tot_inv <- sum(data$Call == "Invalid")
  
    # Percentage positive
    pc_pos <- round((tot_pos * 100 / tot_samples), 2)
    
    # Create table with Cp counts as percentages of total positives
    idexx_table <- data.frame(tot_samples, tot_pos, tot_inv, pc_pos, cp_5, cp_10, cp_15, cp_20, cp_25, cp_30, cp_35, cp_40)
  
  
    # Set table column names
    colnames(idexx_table) <- c("Samples", "Positives", "Invalids", "% Positive", "% 0-5", "% 5-10", "% 10-15", "% 15-20", "% 20-25", "% 25-30", "% 30-35", "% 35-40")
    
    # Handle missing values that result from getting percentage of 0 positives
    idexx_table[idexx_table == "NaN"] <- 0.00
    
    return(idexx_table)
  }
}


## Call the final numbers function to calculate sample numbers and generate the final report table

numbers <- function(path){
  
  # Set the specified path
  suppressWarnings(setwd(path))

  # Run the three numbers functions for each of the three assays; combine the three results into one dataframe
  tot <- rbind(altona_numbers(), euroimmun_numbers(), idexx_numbers())
  
  # Count the total samples, total positives, total invalids, and the percentage positives
  tot_samples <- sum(tot$Samples)
  tot_pos <- sum(tot$Positives)
  tot_inv <- sum(tot$Invalids)
  tot_pc_pos <- round((tot_pos * 100 / tot_samples), 2)
  
  # Total positives with Cp 0 to 5
  cp_5 <- sum(tot$`% 0-5`)
  pc_cp_5 <- round((cp_5 * 100 / tot_pos), 2)
  
  # Total positives with Cp 5 to 10
  cp_10 <- sum(tot$`% 5-10`)
  pc_cp_10 <- round((cp_10 * 100 / tot_pos), 2)
  
  # Total positives with Cp 10 to 15
  cp_15 <- sum(tot$`% 10-15`)
  pc_cp_15 <- round((cp_15 * 100 / tot_pos), 2)
  
  # Total positives with Cp 15 to 20
  cp_20 <- sum(tot$`% 15-20`)
  pc_cp_20 <- round((cp_20 * 100 / tot_pos), 2)
  
  # Total positives with Cp 20 to 25
  cp_25 <- sum(tot$`% 20-25`)
  pc_cp_25 <- round((cp_25 * 100 / tot_pos), 2)
  
  # Total positives with Cp 25 to 30
  cp_30 <- sum(tot$`% 25-30`)
  pc_cp_30 <- round((cp_30 * 100 / tot_pos), 2)
  
  # Total positives with Cp 30 to 35
  cp_35 <- sum(tot$`% 30-35`)
  pc_cp_35 <- round((cp_35 * 100 / tot_pos), 2)
  
  # Total positives with Cp 35 to 40
  cp_40 <- sum(tot$`% 35-40`)
  pc_cp_40 <- round((cp_40 * 100 / tot_pos), 2)
  
  
  # Combine all counts into one dataframe to generate the final numbers report
  final_table <- data.frame(tot_samples, tot_pos, tot_inv, tot_pc_pos, pc_cp_5, pc_cp_10, pc_cp_15, pc_cp_20, pc_cp_25, pc_cp_30, pc_cp_35, pc_cp_40)
  
  # Set column names
  colnames(final_table) <- c("Samples", "Pos", "Inv", "% Pos", "% 0-5", "% 5-10", "% 10-15", "% 15-20", "% 20-25", "% 25-30", "% 30-35", "% 35-40")
      
  # Handle missing values that result from getting percentage of 0 positives
  final_table[final_table == "NaN"] <- 0.00
  
  # Print nicely formatted table using kableExtra
  kable(final_table, row.names = FALSE) %>%
    kable_styling(full_width = F, position = "center")
}
```


### Step 4

This step generates the report. The function is in the form `numbers('path')`, where 'path' is the location of the folder you want to analyse. 

Type in the folder and click the green arrow button to the right.

An example is given below.

```{r, warning = FALSE, echo = TRUE, include = TRUE}
numbers("02 February/02022021/For NVRL Export")
```



#### Important Usage Note

For the sake of convenience, it is not necessary to type the entire folder path; just the path from "01 January" onwards.

Also, Windows separates folder names with a `\`, while RStudio uses a `/` or else it won't work. Please ensure the right character is used. 


#### Troubleshooting

Speak to Conor or Dan with any issues.

